The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}

### 2017-11-24 referee_reply.txt

We thank the referee for a very professional and constructive report. We have
done our best to take the suggestions into consideration and the message of the
paper has been strengthened as a consequence.

Kindest regards

Jonatan Selsing

------------------- Referee Report

The authors present a revised redshift for the short GRB 111117A that will have
a significant impact on the redshift distribution of short GRBs, and potentially
the inferred delay times and detectability of short GRB hosts. This makes the
host of GRB 111117A an outlier in several respects, and a benchmark event, which
naturally merits publication.

While the results are inherently very interesting, the paper suffers from lack
of attention to detail, which are summed up in several major and minor comments
that I have listed below. I feel the authors should address these or explain
their reasoning otherwise before publication.

Major comments: --------------

1) VLT spectrum identification of emission lines: The revised redshift of
z=2.211 relies on the detection of several emission lines in the VLT/X-Shooter
spectrum. Because this is the main point of the paper, the reader really needs
to be convinced that this redshift is correct. While it is clear from Figure 2
that [OIII] 5007 is detected, the determination of the redshift relies on
multiple lines. It is also not obvious from the unbinned spectrum shown in Fig.
3 that the lines are actually detected, because the SED fit line is much
thicker. Thus, the following points should be addressed to make a stronger case
for why the authors think this redshift is correct:

-The authors should show the 2D images (as they have done for [OIII]) for all
lines that they identify.

JS: We agree and we have added a panel showing all the detected lines, which
highlights the certainty of the spectroscopic redshift.

-While it is good to show the unbinned spectrum, I suggest either binning the 1D
spectrum to convince the reader that there are actually emission lines present,
or providing plots that zoom in at the locations of the lines. I suggest
adjusting the plot such that the reader can see the actual data (as currently
one can only see the model due to the thickness of the line).

JS: While we agree that it would be nice if the presence of the lines could be shown
in the 1D spectrum, the lines are too narrow and the S/N of the spectrum too low
for the individual lines to stand out clearly in the full 1D-spectrum,
independently of the binning chosen. We believe that the newly added panel of
the lines in the 2D-spectra convincingly demonstrate the robustness of the
inferred redshift and a corresponding panel for the 1D-lines would not add
valuable information.

-What is the uncertainty in the redshift?

JS: An attempt to attach an uncertainty to the redshift has been added, based on
individual redshift measurement using each of the detected lines.

-The 1D error spectrum should be plotted and the positions of the emission lines
in the 1D spectrum should be labeled.

JS: We agree and this has been included in the spectrum.

2) Deduction of star formation rates:

-When calculating the SFR from H-alpha, it is not clear why this is a lower
limit if the authors have corrected H-alpha for the missing flux.

JS: We have clarified this, which was stated incorrectly in the text. When the
deriving the SFR, we did in fact not correct for the missing flux of H-alpha,
but only used the region free from contamination. This also explains why the
inferred SFR is a limit.

-For both H-alpha and OIII, the fits to the line profiles (after
masking/correcting for the sky lines) should be shown.

JS: Because the profiles are not gaussian or particularly well behaved and given the
low S/N of the spectrum, we have not fitted any of the lines. This is now stated
explicitly in the text. Additionally, OIII was never used to infer the SFR and
this has been cleared up in the text.

-The text currently says, "From the integrated H-alpha line, we estimate SFR =
18 +/- 3 Msun/yr." Is this a typo and H-alpha should really be OIII? If that's
the case, the OIII line is very sensitive to the ionization parameter. The
authors should state what ionization parameter they are assuming, and also
reference any caveats in using OIII as a SFR indicator. The authors should also
make clear how the uncertainty in the OIII SFR was derived.

JS: In the original version of the text, we were referring the h-alpha, although the
text was ambiguous. We have now decided to stick to only the lower limit,
because it was unclear how the systematics of the procedure affected the
uncertainty in the reported error.

-Finally, even though weakly detected, it would be interesting to know if the
derived SFR from H-beta is consistent with the results from H-alpha and OIII.

JS: While certainly interesting, this was also attempted to constrain the reddening
through the use of the Balmer decrement, but unfortunately the large uncertainty
on the H-beta flux prevents obtaining useful information.

3) SED fit: The authors demonstrate a best-fit model of an unreddened galaxy
template with various best-fit parameters for the stellar population properties.
What is the goodness-of-fit parameter used? If it is chi^2 or some other
statistic, the reduced chi^2 (or equivalent) should be stated. The range of
models that were considered (in terms of stellar mass, stellar population age,
and star formation rate), should be included. Was a metallicity assumed, and if
so, what metallicity and why? Finally, how do these inferred properties compare
to previous works?

JS: The SED fit is a template fit, and not a parametric fit. Hence, the number of
"degrees of freedom" is ill-defined, and the reduced chi^2 test does not apply
(e.g. see Andrae et al. 2010 - https://arxiv.org/abs/1012.3754). We have updated
and expanded the text to clarify the procedure. We refer to previous work,
where all the details can be found. In particular we use 4 different input
metallicities, which we have now explicitly stated in the text. We have not
compared the inferred values to previous works, since the previous SED fits are
invalidated by the revised redshift.

4) Photometry: The authors mention that the revised z-band measurement is the
primary reason for why the redshift was previously derived incorrectly by two
different groups. It is interesting and worrisome that the two previous works
(Margutti+ and Sakamoto+) both derived the same "wrong" z-band photometry based
on observations from entirely different facilities (Margutti+ with Gemini, 23.04
+/- 0.18 mag vs. GMG, z=22.9 +/- 0.3 mag).

JS: Indeed the photometric redshift provided by the two previous papers relies on
the identification of the 4000 AA break constrained by the overestimated z-band
flux. We did worry about the consistency of the two previous measurements, both
inconsistent with ours. We do not have access to the GMG data, but we did
download and reduce the Gemini data from which we measure the reported
value/limit. Objects of magnitude z ~ 23.2 are clearly seen and are
significantly brighter than the GRB host galaxy, which is only marginally (if at
all) detected. We did contact both Sakamoto and Margutti and relayed our finding
to them, which they acknowledged, but did not rebut. We do not know what
affected the previous measurements, but we are not in position where we can
answer this. For the GMG data, taken at early times, potentially an afterglow
contribution could influence the measurement as noticed by those authors. We
provide all the reduced imaging data used at
https://github.com/jselsing/GRB111117A, from which the significance of the
detection can be visually confirmed.

-The revised Gemini z-band photometry in Table 1 has an uncertainty of 0.47 mag,
a 2.1-sigma detection, which in most cases would be quoted as an upper limit. Is
there actually visible host emission in the Gemini image, and how was photometry
performed on an object with such low significance? Was the center of the
aperture forced to be at a position where the host is well-detected in a
different band? The authors should state what size aperture they used relative
to FWHM as well as the background region they used, not only for the Gemini
z-band photometry but for all of the photometry listed. In fact, if the Gemini
z-band was quoted properly as a 3- or 5-sigma upper limit, would it still be
consistent with the VLT/FORS2 measurement by the authors?

JS: We agree that the methodology was not adequately explained in the paper and we
have expanded the text regarding the photometric methodology, especially
commenting on the discrepancy between the z-band measurements. For the Gemini
image, we find a limiting 3(5)-sigma upper limit of z > 24.06(23.50) which is
consistent with the FORS measurement within the errors and with our 2-sigma
detection in the GMOS-N image. Note that it is not our intent to claim a
detection in the GMOS-N image as we have already a robust detection in the FORS
image. It is however reassuring that an object of consistent magnitude is not
excluded by the Gemini data. We have added a mosaic of all the available bands,
which shows the significance of the detection in the different bands. A faint
object can be glimpsed at the position of the host in the Gemini panel, visually
consistent with the quote significance of the detection.

-The authors state that the z-band photometry is discrepant from previous works
while other filters are in excellent agreement. A careful comparison between
this work's Table 1, Margutti+ and Sakamoto+ demonstrates that there are
discrepancies not only in Gemini z-band, but in several other bands as well. For
instance, the J-band host measurement reported by the authors is 0.4-5 mag
fainter than that quoted by Sakamoto+ (after correcting for AB to Vega
conversions). The R-band magnitude is 0.4 mag fainter than Sakamoto+, the r-band
magnitude is fainter by 0.2 mag, i-band is 0.25 mag fainter than Margutti+, and
so on. The overall trend is that the host photometry quoted by the authors is
systematically fainter than two previous and separate works. This is worrisome
that two groups got the same wrong answer using different data sets, and should
be addressed by the authors in more detail. Is there something different about
the aperture used or calibration system (SDSS vs. another system, used between
the current work and two previous works that could help explain the
systematically fainter fluxes?

JS: First, please note that values presented in the table are not corrected for the
expected Galactic foreground, contrary to previous works (this amounts to 0.12 /
0.07 / 0.05 in the g/r/i bands). According to what stated in the previous
papers, photometry was calibrated in all cases against SDSS and 2MASS for griz
and JHK filters. In the case of R-band only, Sakamoto et al. use the USNO
catalogue which has a poor photometric accuracy where we use the Lupton color
transformation from SDSS. For the J-band, the discrepancy is 0.4 +- 0.3
(Sakamoto) and 0.5 +- 0.4 (Margutti); for the K-band the discrepancy is 0.35 +-
0.3 (Sakamoto) - all consistent within ~ 1 sigma. For the R-band, our value is
~0.4 +- 0.2 mag fainter than Sakamoto (USNO based calibration), while in the
nearly overlapping r-band we are 0.33 +- 0.15 mag *brighter* (Sakamoto) and 0.15
+- 0.15 mag *fainter* (Margutti). For the i-band we are 0.08 +- 0.3 mag
*brighter* (Sakamoto) and 0.18 +- 0.23 *fainter* (Margutti). The overall picture
is that our values are not systematically fainter (nor brighter) than the
published ones. Most differences are at the 1-sigma level, in some cases our
value is intermediate between the published ones. The numerically large
discrepancy in the infrared stems from the especially large uncertainties, which
are properly accounted for in the fitting process. We agree though that our
statement of an "excellent" agreement was too optimistic, so we have rectified
that statement to saying thar the measurements agree within the errors. Given the
arguments above, we do not feel that a point-by-point discussion of these
discrepancies contributes to the clarity of the message in the paper. As stated
above, though, we have added further explanations on our photometric methodology.

5) NH, AV and Ly-alpha: The high NH, lack of AV, and presence of Ly-alpha is an
interesting aspect of this host. However, the lack of reddening seems more like
an assumption than proven by the data. I have a few questions for clarification:

JS: First we remark that the NH is derived from the X-ray spectrum of the afterglow
and is therefore a sight-line-measurement, where the AV and Lya are integrated
measures. This is particularly relevant given the sizeable offset between the
GRB location and the host center. Therefore there does not need to be a direct
relation between these numbers. The host reddening is a parameter that is fit
for in the SED-fit and thus constrained by the photometry. This is now stated in
the expanded description of the SED fitting.

-The high inferred intrinsic NH is interesting but also matches the evolution of
NH with redshift nicely. There are relations between NH and AV - what is the
derived value for AV and is it consistent with no reddening?

JS: While certainly interesting, the proposed comparison is not straightforward.
Several studies have compared the afterglow-derived extinction with the
afterglow-derived hydrogen equivalent column density. Comparison between the
integrated (luminosity weigthed) host extinction and hydrogen column from X-rays
has not been attempted to our knowledge. In the case of long GRBs, Perley et al.
(2013, ApJ 778 128, their Figure 15) have compared the extinction measured along
the afterglow sightline with the average extinction measured in the host from
SED-fitting, finding consistency between the majority of points within a factor
of a few, but with a large number of outliers. To our knowledge, a corresponding
study has not been carried out for short GRB environments. We have added a small
section discussing this, but we believe this point can be left future works.

-How does the inferred NH compare to previous works from Sakamoto and Margutti?

JS: Because of the higher redshift used, the intrinsic NH derived for our work
is correspondingly higher. We do not feel it is necessary to state this directly in the text.

-A galaxy can still have detectable Ly-alpha in the presence of dust as it
depends on the escape fraction. Are the authors able to make any quantitative
statments on the Lyman-alpha escape fraction, and place a limit on the dust
column?

JS: We have added a section describing the contraints on the Lya escape
fraction and the implications for the dust column. Note, we do not explicitely
calculate a dust column based on the escape fraction, because the dust in the
host is constrained by the photometry.

-Overall, how does the assumption of zero reddening affect the conclusions of
the paper? For instance, if a reddened model is assumed in the SED fit, how
drastically does this change the stellar population parameters?

JS: Zero reddening is not an assumption, but something that is constrained by the
photometry. It is a fit parameter in the SED-fit. We have added caveats to the
SED-fit section.

6) Long versus short GRB classification: The authors claim that GRB 111117A was
not a collapsar, despite the host having a relatively high star formation rate
and potentially high-density local environment. The reader should be convinced
that this is actually the case and some suggestions are below.

-The classification from Bromberg+ 2013 does not seem to match the conclusions
from host galaxy properties (see Berger+ 2014; Fong+ 2013), and is not a very
reliable metric for whether or not an event was a collapsar. While the authors
state that the burst is consistent with zero lag, the authors should state basic
information about the burst, such as duration, hardness, limit on extended
emission, etc to make the case that this event was in fact a short hard GRB.

JS: We can appreciate this point, and have added the basic high-energy properties of
the burst, calculated by previous authors, which clearly marks it as a sGRB.

-The mention of the intermediate class of bursts is interesting, but what is GRB
111117A with respect to this class? Are the host properties of 060121 and 090426
different than those of more definite short GRBs?

JS: Also according to the classification of Horvath et al. 2010, GRB111117A is a
short GRB. This is now stated exclicitely in the text.

7) Implications for delay time distribution: The high redshift of GRB 111117A
indeed has some important implications on the detectability of hosts and
potential delay time distribution. This point of discussion merits a bit more
detail and the following are points to consider.

-The authors state that about half (45-60%) of short GRB hosts if they were at
z=2.211 would have been missed based on luminosity arguments. This is a bit
misleading since the sample of short GRB hosts with redshifts is inherently
incomplete due to Swift detection bias as well as difficulty of following every
short GRB sightline. Do the D'Avanzo+ 2014 and Berger 2014 samples include
elliptical host galaxies where the majority are L* or brighter? If the authors
only include emission line hosts, that actually neglects ~30% of the short GRB
population which come from elliptical galaxies.

JS: The D'Avanzo+ 2014 sample is chosen because it is flux-limited, but otherwise
complete, meaning only unbiasing selection criteria has been imposed on it. It
should therefore correctly reflect the full, underlying sGRB population down to
some sGRB brightness level. The Berger 2014 sample is not complete, which is why
no conclusions are drawn based on this sample. And yes, both samples also
consider the sGRBs originating is elliptical host galaxies. These points have
been clarified a bit in the text.

-Long GRBs originate in star-forming galaxies with rates comparable to that
inferred for the host of GRB 111117A, while long GRB host luminosities are
generally much lower than short GRB hosts. In the scenario that we are missing
many short GRB redshifts because they originate at z~2, why are we routinely
able to detect long GRB hosts to high redshifts, even when they have lower
luminosities?

JS: This is because the large majority of lGRB redshifts (~70 %) come from GRB
afterglow spectroscopy. Only a single sGRB redshift has been found from
afterglow spectroscopy (GRB 130603B). We have added a small section to highlight
the difference in redshift determinations of the two classes of bursts.

-Given the observed short GRB redshift distribution (many at z<1.3, and only a
handful at higher redshifts of z>1.5), it still seems very plausible that the
log-normal delay time distribution is correct. However, if the authors are
correct that the redshift of GRB 111117A is in fact close to the peak of the
short GRB rate and the power-law delay time model is preferred, what is the
implied rate at low redshifts of z<1? Is this consistent with the observed short
GRB rate out to z~1 (which should be fairly spectroscopically complete)? Are
there enough Swift short GRBs without redshift measurements that is consistent
with an explanation that the majority of events actually occur at z~2?

JS: This is a very interesting question and definitely something that merits some
consideration. We have added a section to address this.

Minor comments: ----------------

Abstract: -"between 43 and 71 percent of all sGRB redshifts should be missed" --
what do you mean by all sGRB redshifts? all sGRBs that occur in the Universe,
all of those detected by Swift, or all based on the sample of short GRBs with
known hosts? I think it is the latter, but I would definitely clarify in the
abstract and when you get to this section in the text.

JS: We have added that this is for the Swift-detected bursts, both in the abstract and in the text.

Section 1:

--1st sentence: Define sGRB, NS. 
JS: Done.

--For detection of kilonovae, also cite Berger+ 2013. 
JS: Definitely.

--"duration T_90 < 2" - make sure to qualify with 2 sec
JS: Done.

--Here and throughout, there is a tendency to use sGRB/short GRB and lGRB/long
GRB. I suggest using one or the other throughout the text. 
JS: We have streamline this a bit, but have kept the wording "long and short GRBs"

--"The bulk of these localizations have associated galaxies at relatively low redshifts... it is
arguably biased towards lower redshifts" -- This is not necessarily true, as the
low redshifts could be a natural consequence of relatively long delay times.
JS: We have cautioned the wording.

--"is tracing with host stellar mass rather than star-formation rate" --
Previous analysis of host galaxy properties have actually found that the short
GRB rate likely traces both stellar mass and star formation (see Berger 2014 and
references therein) 
JS: We have changed the wording to reflect this.

--"will provide new a promising ways" -- delete "a"
JS: Done

Section 2: --The authors mention there were 4 separate epochs of spectroscopy -
is any spectral evolution apparent? This would at least be consistent with the
assumption that there is negligible afterglow contribution from epoch 1. 
JS: We do not see any spectral evolution. This is now explicitly stated in the text.

--Since the spectra were taken across different nights and under different seeing
conditions, details of flux calibration for the combined spectrum should be
included. 
JS: We have added a description of this.

--Is it surprising that [OIII] 4959 is not detected, given such a
strong detection of [OIII] 5007? 
JS: [OIII] 4959 is detected, but very close to a sky-line. This has now been added to the text.

--"...because the slit is not aligned along the
long axis of the host, this value is an upper limit" - I believe this should be
a lower limit 
JS: Correct! This has been changed and expanded.

--In calculating beta_OX, state the actual limits that were used
and in what band. Are these the deepest limits available? 
JS: To our knowledge, these are the deepest limits available, yes. The limits have now been explicitly stated.

### 2017-11-24 GRB111117A.tex
Referee commnets final

The host lies in the brighter end of the expected sGRB host brightness
distribution at $z = 2.211$, and is actively forming stars. Using the host as a
benchmark for redshift determination, we find that between 43 and 71 per cent of
all \textit{Swift}-detected sGRB should be missing redshifts due to host
faintness for bursts at $z\sim2$. The high redshift of GRB~111117A is evidence
against a lognormal delay-time model for sGRBs through the predicted redshift
distribution of sGRBs, which is very sensitive to high-$z$ sGRBs.

From the age of the universe at the time of GRB explosion, an initial neutron
star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
progenitor system is a circular pair of inspiralling NSs. This constraint
excludes some of the longest sGRB formation channels for this burst.
}

@@

There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs)
come from the merger of neutron stars (NSs), either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The classification of GRBs in two groups, initially comes from the bimodal
distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
where the duration T$_{90} < 2$ sec has been regarded as the dividing line between long
and short GRBs. Additionally, it has been found for lGRBs that there is a
spectral lag in the arrival-time of photons, with the most energetic ones
arriving first. This lag is consistent with zero for sGRBs
\citep{Norris2006}. Because both populations have continuous, overlapping
distributions in their observables and because telescopes observe in differing
bands, it is difficult to impose a single demarcation criterion between the two
classes. For this reason, the distinction between long and short GRBs is
preferably based on a combination of high-energy properties \citep{Zhang2009,
	Bromberg2012a, Bromberg2013}.

The \textit{Swift} satellite \citep{Gehrels2004} greatly improved the
understanding of sGRB progenitors thanks to its quick localization capability.
The bulk of these localizations have associated galaxies at relatively low
redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, and because
most of these measurements come from the associated hosts, it is potentially
biased towards lower redshifts due to Malmquist bias. The host galaxies of sGRBs
are diverse. They are more massive and less actively star-forming on average
than lGRB hosts \citep{Fong2013b}, while in some cases, no host galaxy can be
identified above detection threshold of deep follow-up observations
\citep{Berger2010a, Tunnicliffe2014}. Together with their position within their
hosts \citep{Fong2013a}, this suggests a progenitor system that can be very long
lived in comparison to lGRBs, with the host stellar mass affecting the sGRB
rates more than star-formation rate (SFR) \citep{Berger2014}.

The electromagnetic signals from sGRBs are a promising channel to accurately
localize NS mergers, which holds the promise for a detection of an associated
gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
of a sGRB and a GW will provide new promising ways to constrain the binary
inclination angle \citep{Arun2014} and measure cosmological distances
\citep{Nissanke2010}.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/OIII_img.pdf}
	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:line}
\end{figure}

@@

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across
three spectroscopic arms. We carried out the bias-correction, flat-fielding,
order tracing, wavelength calibration, rectification, and flux calibration using
the VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run
in physical mode. Because the echelle orders are curved across each detector, a
rectification algorithm is employed which introduces correlations between
neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
maximal resolution. The observations are combined and extracted using scripts
described in Selsing et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
where the full spectral point spread function is modeled across each arm and
used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
slit position is shown in Fig.~\ref{fig:spec_setup}. Each of the epochs are
extracted individually and combined in a weighted fashion where the weights at
each pixel is chosen as median variance spectrum of the region surrounding that
pixel, thus avoiding to base the weight on the pixel variance. An attempt to
correct for slit-losses has been carried out on the combined spectrum based on
the average seeing of the observations. We show the extracted spectrum in
Fig.~\ref{fig:SED}.

@@

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959,
\oiii$\lambda$5007, and \ha. \hb{} and \oiii$\lambda$4959 are both detected at
low significance ($\sim$ 3-$\sigma$). We show \oiii$\lambda$5007 in the insert
in Fig.~\ref{fig:line}.  \ha{} is only visible in the first epoch due to the
$K$-band blocking filter used for the remaining observations. The nebular lines
exhibit a spatial extent of $\sim$ 1\farc5 and show significant velocity
structure along the slit. A drop in the continuum bluewards of the \lya{} line
further supports the inferred redshift. No spectral evolution is observed across
the epochs, supporting the lack of afterglow contribution.

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed
at 21\,000~\AA{} where the night sky is very bright. In addition, several bright
sky-lines are superposed on the line, making an accurate estimate of the
\ha-flux difficult. Due to their velocity structure, the lines exhibit clear
deviations from a Gaussian and given the low S/N of the spectra we do not
attempt any parametric fits. We instead obtain a limit on the SFR by numerically
integrating the part of \ha{} free of contamination and obtain $F_{H \alpha} >
4.1 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After
converting the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial
mass function using \citet{Madau2014}, we derive a limit of $SFR > 7
M_\odot$~yr$^{-1}$. We additionally obtain the \lya{} line flux by numerically integrating
the entire \lya{} line complex and obtain $F_{\mathrm{Ly}\alpha} = 2.0 \pm 0.5
\times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$.

@@

Because the projected distance does not change significantly between $z = 1.3$ and $z =
2.211$, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013}
relating to host offset are unaffected.

>>>>

The host lies in the brighter end of the expected sGRB host brightness
distribution at $z = 2.211$, and is actively forming stars. Using the host as a
benchmark for redshift determination, we find that between 43 and 71 per cent of
all \textit{Swift}-detected sGRBs are too faint at $z\sim2$ to allow for a secure
redshift determination. The high redshift of GRB~111117A is evidence against a
lognormal delay-time model for sGRBs through the predicted redshift distribution
of sGRBs, which is very sensitive to high-$z$ sGRBs.

From the age of the universe at the time of GRB explosion, an initial neutron
star (NS) separation of $a_0 < 3.1~R_\odot$ is required in the case where the
progenitor system is a circular pair of inspiralling NSs. This constraint
excludes some of the longest sGRB formation channels for this burst.
}

@@

There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs) come
from the merger of neutron stars (NSs), either with another NS, or a black hole,
due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. This association was
recently confirmed by the simultaneous and co-spatial detection of gravitational
waves (GWs) from binary neutron star merger and a sGRB
\citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}. The
absence of associated supernovae in deep searches
\citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011} supports this idea and
distinguishes the physical origin of sGRBs from their long-duration
counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The classification of GRBs in two groups initially comes from the bimodal
distribution of burst duration and spectral hardness in the BATSE sample \citep{Kouveliotou1993},
where the duration T$_{90} < 2$ sec has been regarded as the dividing line between long
and short GRBs. Additionally, it has been found for long GRBs (lGRBs) that there is a
spectral lag in the arrival-time of photons, with the most energetic ones
arriving first. This lag is consistent with zero for sGRBs
\citep{Norris2006}. Because both populations have continuous, overlapping
distributions in their observables and because telescopes observe in differing
bands, it is difficult to impose a single demarcation criterion between the two
classes. For this reason, the distinction between long and short GRBs is
preferably based on a combination of the high-energy properties \citep{Zhang2009,
	Kann2011, Bromberg2012a, Bromberg2013}.

The \textit{Swift} satellite \citep{Gehrels2004} greatly improved the
understanding of sGRB progenitors thanks to its quick localization capability.
The bulk of these localizations have associated galaxies at relatively low
redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, but because
most of these measurements come from the associated hosts, it is potentially
biased towards lower redshifts. Additionally, because the \textit{Swift}
sensitivity to sGRB decreases sharply with redshift \citep{Behroozi2014}, the
intrinsic redshift distribution of sGRBs is largely unknown at higher redshifts.

The host galaxies of sGRBs are diverse. They are more massive and less actively
star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases,
no host galaxy can be identified above the detection threshold of deep follow-up
observations \citep{Berger2010a, Tunnicliffe2014}. Together with their position
within their hosts \citep{Fong2013a}, this suggests a progenitor system that can
be very long lived in comparison to lGRBs, with the host stellar mass affecting
the sGRB rates more than star-formation rate (SFR) \citep{Berger2014}.

The electromagnetic signals from sGRBs are a promising channel to accurately
localize NS mergers, which holds the promise for a detection of an associated
gravitational wave (GW) signal \citep{Ghirlanda2016}. This epochal breakthrough
occurred recently when the first NS-NS GW event ever was detected by LIGO/Virgo
(GW 170817) and associated to the weak sGRB 170817A detected by the Fermi and
Integral satellites \citep{LIGOScientificCollaboration2017a, Goldstein2017,
	Savchenko2017}. The simultaneous detection of a sGRB and a GW provides new
promising ways to constrain the binary inclination angle \citep{Arun2014, LIGOScientificCollaboration2017a} and
measure cosmological distances \citep{Nissanke2010, LIGOScientificCollaboration2017c}.

@@

%\begin{figure}
%	\centering
%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
%	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
%	\label{fig:line}
%\end{figure}

\begin{figure*}
	\centering
	\includegraphics[width=18cm]{figures/grid.pdf}
	\caption{2D-images of the emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha. The location of a bright skylines are marked by the blue boxes. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction. For the upper, left panel containing \lya, the systemic redshoft position of \lya is marked by a yellow, vertical, dashed line. The red part of the \oii$\lambda$3727-doublet is affected by atmospheric absorption.}
	\label{fig:line}
\end{figure*}

@@

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, with 5
arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across
three spectroscopic arms. We carried out the bias-correction, flat-fielding,
order tracing, wavelength calibration, rectification, and flux calibration using
the VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run
in physical mode. Because the echelle orders are curved across each detector, a
rectification algorithm is employed which introduces correlations between
neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
maximal resolution. The observations are combined and extracted using scripts
described in Selsing et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
where the full spectral point spread function is modeled across each arm and
used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
slit position is shown in Fig.~\ref{fig:spec_setup}. Each of the epochs are
extracted individually and combined in a weighted fashion where the weights at
each pixel is chosen as median variance spectrum of the region surrounding that
pixel, thus avoiding to base the weight on the pixel variance. Slit-loss
correction has been applied on the combined spectrum based on the average seeing
of the observations. We show the extracted spectrum in Fig.~\ref{fig:SED}.

@@

We determine a redshift of $z = 2.211 \pm 0.001$ from the simultaneous detection
of emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959,
\oiii$\lambda$5007, and \ha. \hb{} and \oiii$\lambda$4959 are both detected at
low significance ($\sim$ 3-$\sigma$). The uncertainty on the redshift is the
standard deviation of independent measurements of the redshift based on the
individual lines. We show \oiii$\lambda$5007 in the insert in
Fig.~\ref{fig:line}.  \ha{} is only visible in the first epoch due to the
$K$-band blocking filter used for the remaining observations. The nebular lines
exhibit a spatial extent of $\sim$ 1\farc5 and show significant velocity
structure along the slit. A drop in the continuum bluewards of the \lya{} line
further supports the inferred redshift. No spectral evolution is observed across
the epochs indicating that there is negligible GRB afterglow contribution to the
spectra.

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed
at 21\,000~\AA{} where the night sky is very bright. In addition, several bright
sky-lines are superposed on the line, making an accurate estimate of the
\ha-flux difficult. Due to their velocity structure, the lines exhibit clear
deviations from a Gaussian and given the low S/N of the spectra we do not
attempt any parametric fits. We instead obtain a limit on the SFR by numerically
integrating the part of \ha{} free of contamination and obtain $F_{H \alpha} >
4.1 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After
converting the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial
mass function using the conversion factor from \citet{Madau2014}, we derive a limit of $SFR > 7
M_\odot$~yr$^{-1}$. We additionally obtain the \lya{} line flux by numerically
integrating the entire \lya{} line complex and obtain $F_{\mathrm{Ly}\alpha} =
2.0 \pm 0.5 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$,
where the error on the flux is found by integrating the associated error
spectrum over the same spectral region.

@@

\citet{Margutti2012} find an projected offset between the host nucleus and the
site of GRB~111117A of 1\farc25 $\pm$ 0\farc20 arcsec and \citet{Sakamoto2013}
find a corresponding offset of 1\farc0 $\pm$ 0\farc20 arcsec. This corresponds
to a projected physical offset at $z = 2.221$ of $10.6 \pm 1.7$ kpc and $8.5 \pm
1.7$ kpc respectively. Because the projected distance does not change
significantly between $z = 1.3$ and $z = 2.211$, all conclusions of
\citet{Margutti2012} and \citet{Sakamoto2013} relating to host offset are
unaffected.
